### 2019-09-04 paper/nature_phys_cover_letter.tex
Added cover letter template for nature physics.

\documentclass{letter}

\usepackage{graphicx}

\begin{document}
\begin{letter}{Nature Physics\\London, UK}

  \opening{Dear Editor:}

  We are submitting a manuscript titled ``The Magnetorotational Instability Prefers Three Dimensions'' for consideration in \emph{Nature Physics}.

\fromsig{\includegraphics[scale=0.5]{jso_sig.pdf}}\\
\fromname{Jeffrey S. Oishi,\\on behalf of the authors}}

\end{letter}
\end{document}